\thetitle

### 2021-05-12 writing/discussion.tex
Add blame trail example to discussion, in new subsection

%% -----------------------------------------------------------------------------

The results of the experiment suggests a number of high-level conclusions about
blame strategies in the gradual typed world.  First, run-time type checks have a
large positive impact, regardless of whether these checks assign blame or throw
plain exceptions.  Second, error messages with blame assignments are more
helpful than those without. The results also indicate, though, that blame is not
critical in a majority of cases, and therefore they suggest investigating
whether blame tracking is worth the performance cost.  Third, the Natural
approach fares better than the Transient approach, but only by a small margin.
Since Natural offers complete and sound path-based blame (i.e. it blames all and only those components
that are relevant to a type impedance mismatch) while Transient offers
incomplete but sound heap-based blame~\citep{gfd-oopsla-2019}, the results call
for a study concerning the relative strengths of the two models of blame.
Fourth, given that Transient's sound but shallow run-time type checks do not
seem to hamper debugging, a version of the Natural semantics of Typed Racket
that disables some of its wrappers may offer an answer to the well-known
performance issues of Natural~\citep{gtnffvf-jfp-2019}.  Fifth, given that both
modes of the Transient rational programmer are equally successful and that
Transient-last produces shorter trails, Reticulated Python could limit the size
of blame sequences to the last few entries to reduce its serious performance
problems (see below).

>>>>

%% -----------------------------------------------------------------------------
\subsection{Contextualizing the results through an example scenario}
\begin{figure}
\input{trail-example-table}
\caption{An example scenario from take5, with every mode's resulting trail.}
\label{fig:example-trails}
\end{figure}

Interpreting the numeric summaries and aggregations described in the last section demands an intuitive understanding of how blame trails look in practice.
Thus, before drawing high level conclusions in the next subsection, this one breaks down one illustrative debugging scenario from the take5 benchmark.
Figure~\ref{fig:example-trails} shows the blame trails produced by every mode of the rational programmer for the debugging scenario in two different ways.
At top left, the trails are depicted diagrammatically as paths through the configuration lattice starting at the root (leftmost) configuration.
Each configuration is represented by a bit string where each digit corresponds to a module in the program and indicates if it is typed.
To contextualize this module information, the module graph of take5 is at top right.
The table at the bottom of the figure expands upon the diagram with the details of every step in each trail.
Specifically, each mode's trail is represented by a row in the table.
The \textbf{Root} column describes the result of running the root configuration in the row's mode.
Using that information, the rational programmer for each mode constructs the successor configuration in the next column (\textbf{Step 1}) if possible.
Repeating the same process from that configuration produces the next column (\textbf{Step 2}).
Finally, the \textbf{Success?} column summarizes whether the trail reaches a type error, and therefore succeeds.
For example, the first row shows that running the root configuration under the Natural-blame mode results in blame on the \texttt{player} module;
typing that module and running again then results in a type error, and hence the trail is successful.
In contrast, the Natural-exceptions row shows that the stack information produced by the root configuration under the Natural-exceptions mode is unhelpful (it only identifies \texttt{main}, which is already typed), and hence the trail immediately gets stuck.
In sum, this example provides a concrete demonstration of how the different modes of the rational programmer behave (and diverge from one another), giving context to the numeric results of the last section.

\subsection{Interpreting the results}
The results of the experiment suggests a number of high-level conclusions about
blame strategies in the gradual typed world.  First, run-time type checks have a
large positive impact, regardless of whether these checks assign blame or throw
plain exceptions.  Second, error messages with blame assignments are more
helpful than those without. The results also indicate, though, that blame is not
critical in a majority of cases, and therefore they suggest investigating
whether blame tracking is worth the performance cost.  Third, the Natural
approach fares better than the Transient approach, but only by a small margin.
Since Natural offers complete and sound path-based blame (i.e. it blames all and only those components
that are relevant to a type impedance mismatch) while Transient offers
incomplete but sound heap-based blame~\citep{gfd-oopsla-2019}, the results call
for a study concerning the relative strengths of the two models of blame.
Fourth, given that Transient's sound but shallow run-time type checks do not
seem to hamper debugging, a version of the Natural semantics of Typed Racket
that disables some of its wrappers may offer an answer to the well-known
performance issues of Natural~\citep{gtnffvf-jfp-2019}.  Fifth, given that both
modes of the Transient rational programmer are equally successful and that
Transient-last produces shorter trails, Reticulated Python could limit the size
of blame sequences to the last few entries to reduce its serious performance
problems (see below).
